---
title: "1.1_Seurat_cluster"
author: "Patrick Maclean"
date: "2024-06-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Seurat)
library(tidyverse)
library(scales)
library(thematic)
```

```{r, include=FALSE}
colour_sequence <- c("#35274A", "#F2300F")

thematic_on(
  bg = "auto", fg = "auto", accent = "auto", font = NA,
  sequential = colour_sequence, qualitative = okabe_ito()
)
show_col(colour_sequence)

theme_set(theme_bw() +
        theme(axis.text = element_text(size = 10)))
```

# Import object
```{r}
ahlers_Seurat <- readRDS(file = "ahlers_Seurat, file = "/well/fullerton/users/rxw025/analyses/skin/scRNA_skin/Processed_data/Ahlers_10xsc/Ahlers_filtered_Seurat.Rds"")
```



# QC

## Find variable features
```{r}
ahlers_Seurat <- FindVariableFeatures(ahlers_Seurat, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(ahlers_Seurat), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(ahlers_Seurat)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 / plot2
```



## Dimensionality reduction
```{r}
# Run PCA on selected 2000 most variable features
ahlers_Seurat <- RunPCA(ahlers_Seurat, features = VariableFeatures(object = ahlers_Seurat))

# Inspect PCs
print(ahlers_Seurat[["pca"]], dims = 1:5, nfeatures = 5)
VizDimLoadings(ahlers_Seurat, dims = 1:2, reduction = "pca")

# Test for elbow on plot
ElbowPlot(ahlers_Seurat, ndims = 40)
```


## Cluster cells
Use KNN algorithm to find cells' neighbours, then draw clusters using Louvain algorithm
```{r}
# Select appropriate number of PCs - go for 20 here
ahlers_Seurat <- FindNeighbors(ahlers_Seurat, dims = 1:20)

# The FindClusters() function implements this procedure, and contains a resolution parameter that sets the ‘granularity’ of the downstream clustering, with increased values leading to a greater number of clusters. We find that setting this parameter between 0.4-1.2 typically returns good results for single-cell datasets of around 3K cells.
# Here I'll use .25 (Ahlers used .2)
ahlers_Seurat <- FindClusters(ahlers_Seurat, resolution = 0.25, cluster.name = 'Unintegrated.clusters')
```

## UMAP
```{r}
ahlers_Seurat <- RunUMAP(ahlers_Seurat, dims = 1:10, , reduction.name = "umap.unintegrated")
```
Visualise
```{r}
DimPlot(ahlers_Seurat,
        reduction = "umap.unintegrated",
        group.by = c('Unintegrated.clusters'),
        ncol = 1) +
          labs(title = "UMAP (no integration)")

DimPlot(ahlers_Seurat,
        reduction = "umap.unintegrated",
        group.by = c('orig.ident'),
        ncol = 1) +
          labs(title = "UMAP (no integration)")

DimPlot(ahlers_Seurat,
        reduction = "umap.unintegrated",
        group.by = c('age'),
        ncol = 1) +
          labs(title = "UMAP (no integration)")
```

# QC with integration
## Integrate
Integrate libraries from the old and young participants
First split the RNA assay by old/young status for integration, then integrate

```{r}
# Older way: FindIntegrationAnchors
FindIntegrationAnchors

# IntegrateData

ahlers_Seurat[["RNA"]] <- split(ahlers_Seurat[["RNA"]], f = ahlers_Seurat$age)


ahlers_Seurat <- IntegrateLayers(object = ahlers_Seurat,
                method = CCAIntegration,
                orig.reduction = "pca",
                new.reduction = "integrated.cca",
                assay = "RNA",
    verbose = T)
```

## Dimensionality reduction

## Cluster
Now find neighbours in the integrated CCA space, and cluster again
```{r}
ahlers_Seurat <- FindNeighbors(ahlers_Seurat, reduction = "integrated.cca", dims = 1:30)
ahlers_Seurat <- FindClusters(ahlers_Seurat, resolution = 0.25, cluster.name = 'Integrated clusters')
```

## UMAP
```{r}
ahlers_Seurat <- RunUMAP(ahlers_Seurat, dims = 1:10, reduction.name = "umap.integrated")
DimPlot(ahlers_Seurat, reduction = "umap", group.by = 'orig.ident')
```

## Compare unintegrated and integrated clusterings
```{r}
DimPlot(ahlers_Seurat, reduction = "umap.unintegrated", group.by = 'orig.ident')
DimPlot(ahlers_Seurat, reduction = "umap.integrated", group.by = 'orig.ident')


# Inspect cluster 5 compared to 0 & 3
cluster5.markers <- FindMarkers(skin.data, ident.1 = 5, ident.2 = c(0, 3))
head(cluster5.markers, n = 5)
```

# Save
```{r}
saveRDS(ahlers_Seurat, file = "/well/fullerton/users/rxw025/analyses/skin/scRNA_skin/Processed_data/Ahlers_10xsc/Ahlers_clustered_Seurat.Rds")
```

